To Ship or Not to Ship: Explict Temporal Ratios Lead to Impatience

### 2015-12-15 paper/paperintro.tex
Start making intro and literature review better

%\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{graphicx}
\usepackage{booktabs,caption,fixltx2e}
%opening
\title{TRaM: Analyzing temporal framing in Intertemporal Choice}
\author{Daniel Wall}

@@

All of us have felt the itch to get the new purchase we made on Amazon as quickly as possible; sometimes leading us choosing expedited shipping, even at considerable cost. 
Generally on Amazon we see shipping times as the number of days until you'll receive your package, e.g.  Receive your package in 2 days. 
However, an identical choice to the one above would be Recieve your package on December 16th. 
\citeA{Read2005} showed that people are more impatient when times are framed as days as compared framed as to dates. 
Accordingly when shopping online delays are occasionally framed as dates.

@@

How the monetary and temporal aspects of an intertemporal choice are framed lead to varying degrees of patience.
Further, recent models of intertemporal choice have posited that people use ratios in intertemporal choices \cite{Read2013, MarzilliEricson2015}. 
%For instance \citeA{Read2013} found that by making ratios in the 
However, to the best of my knowledge, no study has investigated how explicitly framing time as a ratio can alter patience.
Across 3 studies we show that by making ratios explicit people are more impatient in intertemporal choices. 
%Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios.
This result holds for intertemporal choices framed as gains, losses, and in an explict shipping scenario. 
%If the temporal ratio is explicit people are more impatient than when the ratio is implicit.

@@

\section{Intertemporal Choices}

Intertemporal choices in marketing and psychology are generally studied by choices between a smaller sooner and larger later amounts of money \cite{Zauberman2014}.

There are multiple categories of research on these intertemporal choices (see \citeA{Read2013} for an overview) however the current paper belongs to research which investigates how impatience is altered by option frames. 
An example of option framing in intertemporal choice is the hidden zero effect -- when the implicit zeros in intertemporal choices are made explicit people are more patient \cite{Magen2008}. 
For instance when a choice between ``receiving \$100 now'' and ``\$110 in 6 months'' is reframed to be  ``receiving \$100 now \textit{and nothing in six months}'' and ``\$110 in 6 months \textit{and nothing now}'' people are more patient.

Additionally \citeA{Read2013} show that making the ratio between amounts salient affects intertemporal choices. 
Specifically when the magnitude of amounts is small, reframing the amounts as interest rates leads to increased patience. Taking the above example, when reframed as an interest rate, e.g.  a choice between  ``receive \$100 now'' and ``receive 10\% more in six months'' people are more patient. \citeA{Read2013} posit that intrest framed amounts lead to an increased reliance on ratios.

Other work by \citeA{Read2005} shows that how a times are framed in intertemporal choices can alter patience. 
Specifically they find that by reframing the delay above to be a date induces greater patience. 
Taking the above example, people are more patient when reframed it as a choice between  ``receive \$100 now'' and ``receive \$110 on June 15''.
This demonstrates that changing how times are framed can affect patience.

In other work on intertemporal choices, \citeA{MarzilliEricson2015} show that in  standard framed intertemporal choices, people rely on both differences and ratios between both monetary amounts and times.
 Their model -- Intertemporal Choice Heuristic (ITCH) -- does better at predicting out of sample intertemporal choices than classic models of intertemporal choices such as hyperbolic or exponential disounting. 
As seen in Equation~\ref{eq:itch}  the functional form of the itch model with estimated parameters shows that an increased reliance on ratios in time leads to steeper discounting.

@@

This paper combines the work of \citeA{Read2013} -- which shows that framing amounts as ratios alters patience -- and \citeA{MarzilliEricson2015} -- which shows that people rely on differences and ratios in standard intertemporal frame as well as the work  of  \citeA{Read2005} which shows that how times are framed can alter patience. 
Specifically, I focus on how framing times as ratios can decrease patience.
When delays are sufficiently small, and temporal ratios are made explicit, our attention will be focused on how much \textit{longer} we will wait. 
This focus on how much longer we will wait will lead to decreased patience when times are framed as ratios.

I operationalize this prediction across three studies.
Study 1 investigates how making temporal ratios explicit in intertemporal choices leads to greater impatience.

Study 2 extends the results of Study 1 to losses. 
Although most intertemporal choice studies use gains, a large number of real world intertemporal choices involve losses \cite{Hardisty2013}. 
For instance deciding to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
Intertemporal choices framed as losses take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
Further most studies investigating framing effects on intertemporal choices are restricted to the domain of gains. 
Study 2 investigates how framing times as ratios leads to decreased patience in losses.

Study 3 extends the results from Study 1 and Study 2 to intertemporal choices explicitly framed as shipping scenarios.
Shipping scenarios are displayed as a choice between ``Receive your items in 2 days and pay \$8'' and ``Receive your items in 10 days and pay \$4''
Although choosing a shipping speed is similar to the intertemporal choices there is a subtle, yet possibly significant difference between the two.
Shipping decisions are paying more now to receive an item sooner whereas the intertemporal choices shown in study 2 are paying something smaller now to avoid paying something larger later.

@@

@@

\subsection{Method}
Two hundred participants from Amazon Mechanical turk we recruited to take part in a ``Decision Making Survey''. 
Although the validity of Mechanical Turk has been called into question, \citeA{Paolacci2010} finds that classic effects in decision making replicate. 
Furhter, participants who have less than 50 HITs accepted or  who have less than 95\% of their HITs accepted were not allowed to participate in the study.

@@

Insert a qualtrics picture here.

@@

After completeing their choices participants completed demographic information.

>>>>

%\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{graphicx}
\usepackage{booktabs,caption,fixltx2e}
%opening
\title{To Ship or Not to Ship: Explict Temporal Ratios Lead to Impatience}
\author{Daniel Wall}

@@

All of us have felt the itch to get the new purchase we made on Amazon as quickly as possible; sometimes leading us choosing expedited shipping, even at considerable cost. 
Generally on Amazon we see shipping times as the number of days until you will receive your package, e.g.  ``Receive your package in 2 days''. 
The shipping time could also be described as ``Receive your package on December 17th''. 
This reframing of delays -- 2 days -- into dates -- December 17th -- has been shown to increase patience \cite{Read2005}.
Occasionally Amazon frames shipping times as dates, possibly to induce customers to be more patient and receive their items later.

@@

How the monetary and temporal aspects of an intertemporal choice are framed lead to varying degrees of patience.
Further, recent models of intertemporal choice have posited that people use ratios in intertemporal choices \cite{Read2013, MarzilliEricson2015}. 
%For instance \citeA{Read2013} found that by making ratios in the 
However, to the best of my knowledge, no study has investigated how explicitly framing time as a ratio can alter patience.
Across 3 studies I show that by making ratios explicit people are more impatient in intertemporal choices. 
%Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios.
This result holds for intertemporal choices framed as gains, losses, and in an explicit shipping scenario. 
%If the temporal ratio is explicit people are more impatient than when the ratio is implicit.

@@

\section{Literature Review}

Intertemporal preferences measured in marketing and psychology are generally studied by choices between a smaller sooner and larger later amounts of money \cite{Zauberman2014}. 
There are four general categories of research on  intertemporal choices (see \citeA{Read2013} for an overview).
The first focuses on how magnitude of both time and money effects patience \cite{Scholten2006, Thaler1981}
 the current paper belongs to research investigating how patience is altered by option frames. 
 The second stream focuses on how choices relate to real world behavior.
 Individual differences in imputed discount functions predict real world phenomenon such as credit card debt,  mortgage choice, and smoking \cite{MacKillop2011, Meier2009, Johnson2011}. 
 The third stream focuses on how emotional states can alter intertmporal preferences, e.g. being seeing freshly baked bread \cite{Li2008}. 
 The fourth stream investigates how the framing of options affects patience.

An example of option framing in intertemporal choice is the hidden zero effect -- when the implicit zeros in intertemporal choices are made explicit people are more patient \cite{Magen2008}. 
For instance when a choice between ``receiving \$100 now'' and ``\$110 in 6 months'' is reframed to be  ``receiving \$100 now \textit{and nothing in six months}'' and ``\$110 in 6 months \textit{and nothing now}'' people are more patient.

Additionally \citeA{Read2013} show that making the ratio between amounts salient affects intertemporal choices. 
Specifically when the magnitude of amounts is small, reframing the amounts as interest rates leads to increased patience. Taking the above example, when reframed as an interest rate, e.g.  a choice between  ``receive \$100 now'' and ``receive 10\% more in six months'' people are more patient. \citeA{Read2013} posit that interest framed amounts lead to an increased reliance on ratios of monetary amounts.

Beyond amounts, times can also be reframed.
The most famous example of temporal reframing is  Lincoln's Gettysburg Address which made 87 years into the ``Four score and seven years ago''.
\citeA{Read2005} shows that how a times are framed in intertemporal choices can alter patience. 
Taking the above example, people are more patient when reframed it as a choice between  ``receive \$100 now'' and ``receive \$110 on June 15'' -- this is known as the delay/date effect. 
The delay/date effect demonstrates that perceived temporal distance is dependent on how delays are framed.

In other work on intertemporal choices, \citeA{MarzilliEricson2015} show that in intertemporal choices framed as ``receiving \$100 now'' and ``\$110 in 6 months'', people rely on both differences and ratios of the  attributes of intertemporal choices -- monetary amounts and times.
Their model -- the Intertemporal Choice Heuristic (ITCH) -- does better at predicting intertemporal choices than classic models, such as hyperbolic or exponential discounting. 
As seen in Equation~\ref{eq:itch}  the functional form of the ITCH model with estimated parameters shows that an increased reliance on ratios in time leads to steeper discounting.

@@

This work aims to demonstrate that the fourth research stream -- that patience is affected by how options are framed -- can affect real world phenomenon.
Specifically, this paper combines the work of \citeA{Read2013} -- which shows that framing amounts as ratios alters patience -- and \citeA{MarzilliEricson2015} -- which shows that people rely on differences and ratios in standard intertemporal frame as well as the work  of  \citeA{Read2005} which shows that how times are framed can alter patience. 
Specifically, I focus on how framing times as ratios can decrease patience.
When delays are sufficiently small, and when temporal ratios are made explicit, our attention will be focused on the ratio of how much \textit{longer} we will wait. 
For instance reframing a choice between ``receiving \$100  in 1 day'' and ``\$110 in 10 days'' to being a choice between ``receiving \$100  in 1 day'' and ``\$110 in 10 \textit{times as long}'', will focus participants on how much longer they will wait for \$110 as opposed to receiving the \$100 today.
This increased focus on how long they will wait, leads to decreased patience when times are framed as ratios.

@@

I test these hypotheses across three studies.
Study 1 investigates Hypothesis 1 by making temporal ratios explicit in intertemporal.

Study 2 extends the results of Study 1 to losses. 
Although most intertemporal choice studies use gains, a large number of real world intertemporal choices involve losses \cite{Hardisty2013}. 
For instance waiting to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
Intertemporal choices framed as lossed measured in the lab take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
Further most studies investigating framing effects on intertemporal choices are restricted to the domain of gains. 
Study 2 investigates how framing times as ratios leads to decreased patience in losses.

Study 3 extends the results from Study 1 and Study 2 to intertemporal choices explicitly framed as shipping scenarios.
Shipping scenarios are typically displayed as a choice between ``Receive your items in 2 days and pay \$8'' and ``Receive your items in 10 days and pay \$4''
Although choosing a shipping speed is similar to the previous intertemporal choices there is a subtle, yet possibly significant difference between the two.
Shipping decisions can be thought of as paying more now to receive an item sooner whereas the intertemporal choices shown in Study 2 can be thought of as paying something smaller now to avoid paying something larger later.

@@

\subsection{Method}
Two hundred participants from Amazon Mechanical turk we recruited to take part in a ``Decision Making Survey''. 
Although the validity of Mechanical Turk has been called into question, \citeA{Paolacci2010} finds that classic effects in decision making replicate. 
Further, participants who have less than 50 HITs accepted or  who have less than 95\% of their HITs accepted were not allowed to participate in the study. 
Additionally only participants who passed an attention check were included in the final analysis.
These precautions were taken to ensure that participants completed

@@

Insert a Qualtrics picture here.

@@

After completing their choices participants completed demographic information.
